---
title: "size"
output: html_document
date: "2023-12-17"
---
******

GOAL : **third pre-treatment with Savitzky Golay rchemo and PLSDA**

*****


```{r project_organization}
project.dir <- "C:/Users/Utilisateur/Snails_Project"
data.dir <- paste0(project.dir, "/Temoins")
doc.dir <- paste0(project.dir,"/Code_Pre_traitement")
```

```{r setup}
knitr::opts_chunk$set(fig.path=project.dir, 
                      echo = TRUE, 
                      message=FALSE, 
                      warning=FALSE)
```

# Recommanded R packages

```{r libraries}

#----------Importation and data management
  library("tidyverse") # data management
  library("readxl")
#-----------------------------------------------------

#----------Visualization and pre-treatment for functional data
  library("hyperSpec") # Spectral data management
  library("prospectr") # Spectral data pre-treatment
  library("rchemo")    # all inclusive
  library(fda.usc)   # functional data management
#-----------------------------------------------------

#---------- Machine Learning
  library("randomForest")# randomForest
  library("MLmetrics")  # Multiclasssummary for caret.
  library("caret")       # machine learning
#-----------------------------------------------------
#---------- Other tools for visualization
  library(MASS)       # basic statistics
  library(gridExtra)  # combining ggplot graphics
  library(cowplot)    # combining ggplot graphics
```

# Data Inspection

## Data importation

```{r importation}
  spectra <- read.table(paste(data.dir,"spectra.csv",sep="/"), sep=",", h=T, row=1)
  class <- read.table(paste(data.dir,"classes.csv",sep="/"), sep=",", h=T, row=1 )
```

## Creation of a vector of the wavelength values

```{r wave}
# Creation of a vector of wavelength values
wavelength.vector <- as.numeric(gsub(pattern="X", 
                                     replacement="",
                                     x=colnames(spectra)[1:216]))

# Convertir les noms de colonnes en valeurs numériques
new_column_names <- as.numeric(gsub(pattern = "X", replacement = "", x = colnames(spectra)[1:216]))

# Changer les noms de colonnes dans l'objet spectra
colnames(spectra)[1:216] <- new_column_names
```


# Data representation

## Groups or classes

```{r group}
spectra_class = spectra %>%
  mutate (group = as.factor(class$num_class))%>%
  mutate (size = as.factor(class$Size))
spectra_class=spectra_class[complete.cases(spectra_class),]
#on créer une matrice spectra_class avec les spectres plus les informations sur la diet en dernière colonne et on en retire les lignes sans informations.
```

## Graphic of the spectral dataa

```{r graprchemo, fig.width=12, fig.eight=4}
# From package rchemo

plotsp(spectra_class[,1:216], 
       col = ifelse(spectra_class$group == 2 ,"green","brown"),
       main = ("Datas from python script"),
      xlab="",ylab="Reflectance")

# Filter rows where spectra_class$group == 2
subset_data_Living <- apply(spectra_class[spectra_class$group == 2,-c(217,218) ], MARGIN = 2, mean)
# Filter rows where spectra_class$group == 1
subset_data_Dead <- apply(spectra_class[spectra_class$group == 1,-c(217,218) ], MARGIN = 2, mean)

# Create a matrix with data from subset_data
subset_data_diet <- matrix(c(subset_data_Living, subset_data_Dead), nrow = 2, byrow = TRUE)
# Name the columns with the same names as spectra
colnames(subset_data_diet) <- colnames(spectra)
# Name the rows with the names of the subset_data
rownames(subset_data_diet) <- c("Living", "Dead")

# Plot the subset of data using plotsp
plotsp(subset_data_diet, lwd = 2,
       col = c("green","brown"),
       main = "Datas from python script, Diet",
       xlab = "Wavelengths", ylab = "Reflectance",Legend=TRUE)

# Add legend
legend("topright", legend = c("Living", "Dead"), col = c("green", "brown"), lty = 1,lwd = 2)

# Filter rows where spectra_class$size == 1
subset_data_large <- apply(spectra_class[spectra_class$size == 1,-c(217,218) ], MARGIN = 2, mean)

# Filter rows where spectra_class$size == 2
subset_data_medium <- apply(spectra_class[spectra_class$size == 2,-c(217,218) ], MARGIN = 2, mean)

# Filter rows where spectra_class$size == 3
subset_data_small <- apply(spectra_class[spectra_class$size == 3,-c(217,218) ], MARGIN = 2, mean)

# Create a matrix with data from subset_data
subset_data_size <- matrix(c(subset_data_large, subset_data_medium, subset_data_small), nrow = 3, byrow = TRUE)

# Name the columns with the same names as spectra
colnames(subset_data_size) <- colnames(spectra)

# Name the rows with the names of the subset_data
rownames(subset_data_size) <- c("Large", "Medium", "Small")

# Plot the subset of data using plotsp
plotsp(subset_data_size, lwd = 2,
       col = c("red","violet","orange"),
       main = "Datas from python script, size",
       xlab = "Wavelengths", ylab = "Reflectance")
# Add legend
legend("topright", legend = c("Large", "Medium","Small"), col = c("red","violet","orange"), lty = 1,lwd = 2)
```


# Pre-treatment of spectral data

## Basic pre-treatment as centering and normalisation from `prospectr` or `rchemo`
```{r basic, fig.width=12, fig.height=4}

# the standard normal variate transformation (rchemo)
# ?snv to have a look at the help
plotsp(snv(spectra_class[,(1:216)]), 
       col = ifelse(spectra_class$group == 2 ,"green","brown"),
        main=("SNV transformation"),
        xlab="",ylab="Spectral data")
```

## Derivatives from `rchemo`

```{r deriv, fig.width=12, fig.height=4}
# p = polynomial order 
# w = window size (must be odd) 
# m = m-th derivative (0 = smoothing) 

sg.rchemo <- rchemo::savgol(spectra_class[,(1:216)], p = 2, n = 5, m = 1)

plotsp(sg.rchemo, 
       col = ifelse(spectra_class$size == 1 ,"brown",ifelse(spectra_class$size == 2,"red","orange")),
       main = "Savitzky Golay rchemo",
       xlab="",ylab="Spectral data")
par(mfrow = c(1, 1))
```

#Save

```{r}
save_path <- paste0(data.dir,"/spectra_correction.csv")
matable <- sg.rchemo %>%
  as.data.frame()%>% 
  mutate(group=spectra_class$group)%>% 
  mutate (size = spectra_class$size)
write.csv(matable, save_path)
```

## Data Importation and Management


## PLS-DA
The usual tool for spectral data is PLS, here in the Discriminant Analysis version to perform supervised classification. In this use case, we used the R package **mixOmics** from the **Bioconductor** platform instead of the **CRAN** platform. To install this package on your computer, do the following steps:

```{r installmixO,echo=TRUE,eval=FALSE,message=FALSE, warning=FALSE}
  if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

  BiocManager::install("mixOmics")  
```

```{r libmixO}
  library(mixOmics) # PLS derived methods for Omics Data
```

### learning from the training datasets of Ixos and Primadur


**Step 1 : Partition of the dataset `sp` into calibration and validation subsets**

```{r sampling}
  # Sampling
  size = round(nrow(matable)/3)
  id_val=sample(1:nrow(matable))[1:size]

  # Calibration and validation datasets
  matable_val=matable[id_val,-c(217,218)]; matable_cal=matable[-id_val,-c(217,218)]
  variety_val= matable$size[id_val]; variety_cal=matable$size[-id_val]
```

**Step 2 : Calibration by hand**
Run PLSDA with different parameter (`ncomp`) values to calibrate the  number of components of the PLSDA.

```{r calibplsda}
  ## PLSDA
# Initialization
  ncmax=20 #Fixe le nombre maximal de composantes principales à considérer dans la boucle à 10.
  perok=vector(mode='numeric',length=ncmax) # Initialise un vecteur perok pour stocker les performances du modèle pour différentes valeurs de composantes principales.
#

# Loop
  for (i in 1:ncmax) {
    # Model Estimation
    rplsda=plsda(matable_cal, variety_cal,ncomp=i)
    # Model performance
    pred=predict(rplsda ,matable_val)
    # Contingency table for 
    t=table(as.factor(pred$class$mahalanobis.dist[,i]),variety_val)
    # error rate
    perok[i]=1-sum(diag(as.matrix(t)))/size
  }
# Performance results
  perok 
```
**Step 2bis : Calibration from mixOmics functions**

click on  [mixOmics plsda vignette](https://bioconductor.org/packages/release/bioc/vignettes/mixOmics/inst/doc/vignette.html#pls-da-on-the-srbct-case-study-05)

To set a more robust calibration, the cross-validation is repeated 20 times for each component value between 1 and `ncomp` in the `perf` function. 

```{r plsda,fig.width=12, fig.height=4}
plsda.res <- plsda(X=matable_cal, Y=variety_cal, ncomp = 9)

perf.plsda.res <- perf(plsda.res, validation = 'Mfold', folds = 5, 
                  criterion = "all",
                  progressBar = TRUE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda.res, sd = TRUE, legend.position = 'horizontal')
```

```{r perfplsda}
perf.plsda.res$choice.ncomp

# max.dist : distance maximale entre les moyennes des classes
# centroids.dist : distance entre les centroïdes des classes. Les centroïdes sont les moyennes des observations appartenant à chaque classe
# mahalanobis.dist : mesure de la distance entre un point et un ensemble de points, en tenant compte de la covariance entre les variables
```

**Estimation of the model**
```{r plsdafinal}
# Estimation of the model from the whole training dataset
# using perf.plsda.res$choice.ncomp for Mahalanobis distance
  rplsda_final=plsda(matable_cal, variety_cal, ncomp=8)
# Performance
  perf.plsda.final <- perf(rplsda_final, validation = 'Mfold', folds = 5, 
                  criterion = "all",
                  progressBar = TRUE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50
  
 # Calcul de la moyenne le long de l'axe 2 pour obtenir une matrice 2D
perf.plsda.final_class_mahalanobis.dist_numeric <- apply(perf.plsda.final$class$mahalanobis.dist, c(1, 3), as.numeric)
perf.plsda.final_class_mahalanobis.dist_2D_moyenne <- apply(perf.plsda.final_class_mahalanobis.dist_numeric, c(2, 3), mean)
  perf.plsda.final_class_mahalanobis.dist_2D_moyenne <- round(perf.plsda.final_class_mahalanobis.dist_2D_moyenne)
perf.plsda.final_class_mahalanobis.dist_2D_moyenne <- as.factor(perf.plsda.final_class_mahalanobis.dist_2D_moyenne[,1])

conf_matrix <- confusionMatrix(perf.plsda.final_class_mahalanobis.dist_2D_moyenne, reference = variety_cal)

# Afficher la matrice de confusion
print(conf_matrix)

# Afficher l'accuracy
accuracy <- conf_matrix$overall["Accuracy"]
cat("Accuracy:", accuracy, "\n")

```


**Loadings**
```{r loading,fig.width=12, fig.height=4}
rplsda_final
#plotLoadings(rplsda_final,comp=1)
plot(colnames(sg.rchemo), rplsda_final$loadings$X[,1],
     type="l",lty=1, 
     xlab="Wavelength",
     ylab="Loadings", 
     main="1st component")
abline(h=0,col="red", lty=2)
plot(colnames(sg.rchemo), rplsda_final$loadings$X[,2],
     type="l",lty=1, 
     xlab="Wavelength",
     ylab="Loadings", 
     main="2nd component")
abline(h=0,col="red", lty=2)
plot(colnames(sg.rchemo), rplsda_final$loadings$X[,3],
     type="l",lty=1, 
     xlab="Wavelength",
     ylab="Loadings", 
     main="3rd component")
abline(h=0,col="red", lty=2)
```

```{r,fig.width=11}
vip_scores <- rplsda_final$loadings$X

# Get loadings
loadings_matrix <- rplsda_final$loadings$X

# Calculate the mean of loadings for each variable
mean_loadings <- rowMeans(loadings_matrix)

# Extract VIP scores for all components
for (i in 1:3) {
  vip_scores_componenti <- vip_scores[, i]
  
  # Calculate the mean of loadings for the current component
  mean_loading <- mean(vip_scores_componenti)
  # Select wavelengths with loadings above the mean
  selected_wavelengths <- names(vip_scores_componenti[vip_scores_componenti > mean_loading])

  ## Optionally, plot VIP scores for visualization

  # Calculate the difference from the mean to adjust the position of the   bars
  difference_from_mean <- vip_scores_componenti - mean_loading
  
  # Extract the response variable (e.g., group)
  response_variable <- matable$group
  # Store results
  results <- data.frame(Wavelength = character(), P_Value = numeric(), stringsAsFactors = FALSE)
    for (wavelength in selected_wavelengths) {
      predictor_variable <- matable[, wavelength]
    
    # Perform ANOVA
      anova_result <- aov(predictor_variable ~ response_variable)
    
    # Extract relevant information from ANOVA result (e.g., p-value)
      p_value <- summary(anova_result)[[1]]$`Pr(>F)`[1]
    
    # Store the results
      results <- rbind(results, data.frame(Wavelength = wavelength, P_Value = p_value))
    
    # Print or store the results as needed
    cat("Wavelength:", wavelength, " | P-Value:", p_value, "\n")
  }
  # Filter the results based on p-value
  significant_results <- results[results$P_Value < 0.05, ]
  
# Optionally, plot barplot with significant wavelengths in red
  barplot(difference_from_mean,
          col = ifelse(colnames(matable) %in% significant_results[,1], "red", "blue"),
          main = paste("ANOVA p_value for Component",i),
          xlab = "Wavelengths",
          ylab = "VIP Scores",
          ylim = c(min(difference_from_mean), max(difference_from_mean)),
          # Set ylim based on the range of adjusted VIP scores
    )
}

vip_scores <- rplsda_final$loadings$X

# Get loadings
loadings_matrix <- rplsda_final$loadings$X

# Calculate the mean of loadings for each variable
mean_loadings <- rowMeans(loadings_matrix)

# Initialisation d'une liste pour stocker les résultats de chaque composant
all_results <- list()

# Extract VIP scores for all components
for (i in 1:3) {
  vip_scores_componenti <- vip_scores[, i]
  
  # Calculate the mean of loadings for the current component
  mean_loading <- mean(vip_scores_componenti)
  
  # Select wavelengths with loadings above the mean
  selected_wavelengths <- names(vip_scores_componenti[vip_scores_componenti > mean_loading])
  # Afficher la matrice finale
  save_path <- paste0(data.dir,"/selected_wavelengths",i,".csv")
  write.csv(selected_wavelengths,save_path)
  # Calculate the difference from the mean to adjust the position of the bars
  difference_from_mean <- vip_scores_componenti - mean_loading
  
  # Extract the response variable (e.g., group)
  response_variable <- matable$group
  
  # Initialisation d'un data frame pour stocker les résultats du composant actuel
  results <- data.frame(Wavelength = character(), P_Value = numeric(), stringsAsFactors = FALSE)
  
  for (wavelength in selected_wavelengths) {
    predictor_variable <- matable[, wavelength]
    
    # Perform ANOVA
    anova_result <- aov(predictor_variable ~ response_variable)
    
    # Extract relevant information from ANOVA result (e.g., p-value)
    p_value <- summary(anova_result)[[1]]$`Pr(>F)`[1]
    
    # Store the results
    results <- rbind(results, data.frame(Wavelength = wavelength, P_Value = p_value))
    
    # Print or store the results as needed
    cat("Wavelength:", wavelength, " | P-Value:", p_value, "\n")
  }
  
  # Filter the results based on p-value
  significant_results <- results[results$P_Value < 0.05, ]
  
  # Stocker les résultats dans la liste
  all_results[[paste("Component", i)]] <- significant_results
}

# Combinez les résultats dans une matrice
final_matrix <- do.call(rbind, all_results)



```

**Explained Variance percentages** (amount of variance explained per component (note that contrary to PCA, this amount may not decrease as the aim of the method is not to maximize the variance, but the covariance between X and the dummy matrix Y).

```{r percvar}
lapply(rplsda_final$prop_expl_var,round,2)
```

```{r plotindiv,fig.width=12, fig.height=4}
# Identify common indices between matable_cal and matable
common_indices <- intersect(rownames(matable_cal), rownames(matable))

# Add the column to matable_cal
matable_cal_group_size <- cbind(matable_cal, group= matable[common_indices, 217],size = matable[common_indices,218])
plotIndiv(rplsda_final,comp=c(1,2),ellipse = TRUE,
          group = rplsda_final$Y,pch=as.numeric(matable_cal_group_size[,217]),
          pch.levels= ifelse(matable_cal_group_size[,217]==1,"Dead","Living") ,legend=TRUE,col=c("red","green","blue"))


```


### Variety identification in the mixture dataset
```{r predplsda, fig.width=12, fig.height=4}
matrix_wild <- read.table(paste(project.dir,"Sauvages/spectra.csv",sep="/"), sep=",", h=T, row=1) 
wild_classes <- read.table(paste(project.dir,"Sauvages/classes.csv",sep="/"), sep=",", h=T, row=1) 
# Creation of a vector of wavelength values
wavelength.vector <- as.numeric(gsub(pattern="X", 
                                     replacement="",
                                     x=colnames(matrix_wild)[1:216]))

# Convertir les noms de colonnes en valeurs numériques
new_column_names <- as.numeric(gsub(pattern = "X", replacement = "", x = colnames(matrix_wild)[1:216]))

# Changer les noms de colonnes dans l'objet spectra
colnames(matrix_wild)[1:216] <- new_column_names

#wild groups prediction
pred_wild=predict(rplsda_final,matrix_wild)

# Assuming you want the coordinates for components 1 and 2
predicted_coordinates <- predict(rplsda_final, newdata = matrix_wild, comp = c(1, 2))

# Now, you can use the predicted_coordinates to plot individuals on a plane
matrix_wild = matrix_wild %>%
  mutate (group = as.factor(wild_classes$num_class))%>%
  mutate (size = as.factor(wild_classes$Size))

control <- plotIndiv(rplsda_final,comp=c(1,2),ellipse = TRUE,
          group = rplsda_final$Y,pch=as.numeric(matable_cal_size[,217]),
          pch.levels= ifelse(matable_cal_size[,217]==1,"large",ifelse(matable_cal_size[,217]==2,"medium","small")) ,legend=TRUE,col=c("brown","green"))

# Add points to the existing mixOmics plot
add.points(control, comp = c(1, 2), x = pred_wild$predict[,,c(1,2)], legend = FALSE, col = "blue", pch = as.numeric(matrix_wild[,218]))
```

```{r}
myresu<-cbind.data.frame(pred_matable$class$mahalanobis,as.numeric(matable.rf$aggregate))
colnames(myresu)<-c(paste0("plsda",seq(1,ncol(pred_matable$class$mahalanobis))),
                      "RF")
head(myresu)
```